Backpropagating Gravitational wave events

### 2022-03-17 src/ms.tex
Update discussion

The full posterior is available in appendix \kw{Or we should w=put a link to the dataset?}

@@

\begin{figure}
\includegraphics[width=0.49\textwidth]{static/GW150914_m1m2alpha.png}
\caption{}
\label{fig:GW150914_M1_M2_q3}
\end{figure}

The most important point in this paper is illustrated in figure \ref{fig:GW150914_M1_M2_q3}.
We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $q_3$,
this means if we fixed the value of $q_3$ instead of allowing it to jointly vary with all other parameters,
we would miss the samples that do not match the chosen value of $q_3$.
By folding in the hyper-parameters into the root-finding algorithm,
we can capture the correlation between 
\kw{Dartboard comparsion} Comparing our result with 
\kw{Add figure showing hyper-parameters are important to recovering the progenitor parameters correctly.}

@@

To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
We "back propagate" the posterior in the observable space to the evolutionary parameter space,
thus allowing us to study the correlation between progenitor parameters and hyper-parameters.
We demonstrate the capability of this method with two events, GW150914 and GW191129, while employing the binary population synthesis model \texttt{COSMIC}.
We 
We hope this letter would motivate the community to build the relevant infrastructure mentioned above,
so in the near future we can explore the full parameter space of binary evolution models with the next-generation GW detectors network.

Interpretation:
\begin{enumerate}
\item Highlight the fact that we turn hyper-parameters into parameters.
\item Discuss this is need for high dimensional exploration.
\end{enumerate}

Caveats: 
\begin{enumerate}
\item Discuss events that cannot be complete transferred backward. From BSE angle and from LIGO posterior angle. Failure mode in mass.
\item We only assume BSE
\item Caveat on marginalizing over random samples.
\item Prior need to be applied after one translate the posterior, i.e. it is a post-processing step now.
\item How next generation "Progenitor synthesis" code should be built.
\item Rate
\item Failure studies go into appendix
\end{enumerate}

>>>>

The full posterior is available in appendix \kw{Or we should put a link to the dataset?}

@@

\begin{figure}
\includegraphics[width=0.49\textwidth]{static/GW150914_m1m2alpha.png}
\caption{}
\label{fig:GW150914_m1_m2_alpha}
\end{figure}

The most important point in this paper is illustrated in figure \ref{fig:GW150914_m1_m2_alpha}.
We can see the posterior samples in $M_1-M_2$ space are spatially correlated with $\alpha$,
this means if we fixed the value of $\alpha$ instead of allowing it to jointly vary with all other parameters,
we would miss the samples that do not match the chosen value of $\alpha$.
By folding in the hyper-parameters into the root-finding algorithm,
we can capture the correlation between 
\kw{Dartboard comparsion} Comparing our result with 
\kw{Add figure showing hyper-parameters are important to recovering the progenitor parameters correctly.}

@@

To summarize, we propose a novel method to recover the posterior samples in the evolutionary parameter space for each GW event.
We point out hyper-parameters in the usual population synthesis simulation context are not actually parameters related to the population,
but parameters about the evolutionary function.
This means the binary evolution functions can be constrained on an individual event basis.
We "back propagate" the posterior in the observable space to the evolutionary parameter space,
thus allowing us to study hyper-parameters and its correlation with progenitor parameters in a data-driven manner.
Our method makes less assumption than the traditional forward modelling approach,
which often fix the hyper-parameters across the entire population.
Since we are not limited to the fix hyper-parameters assumption, we can explore the behavior of the hyper-parameters across the population much more efficiently.
While our work lay down a data analysis pathway to understand the population of GW events,
no physics can be learned without a comprehensive complimentary physical model.
We hope this letter would motivate the community to build the next generation population synthesis codes that have the following properties:
first, they should have explicit control over the random variables so marginalizing over random variables can be done more precisely;
second, they should be as automatically differentiable as much as possible so exploring the evolutionary parameter space is efficient.
Combining this work and the next generation population synthesis code,
we can explore the full parameter space of binary evolution models with the next-generation GW detectors network in the near future.

% Interpretation:
% \begin{enumerate}
% \item Highlight the fact that we turn hyper-parameters into parameters.
% \item Discuss this is need for high dimensional exploration.
% \end{enumerate}

% Caveats: 
% \begin{enumerate}
% \item Discuss events that cannot be complete transferred backward. From BSE angle and from LIGO posterior angle. Failure mode in mass.
% \item We only assume BSE
% \item Caveat on marginalizing over random samples.
% \item Prior need to be applied after one translate the posterior, i.e. it is a post-processing step now.
% \item How next generation "Progenitor synthesis" code should be built.
% \item Rate
% \item Failure studies go into appendix
% \end{enumerate}
